%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QC-Amp: Quantum Computation of QCD Colour Factors
% PhD Thesis Chapter / Technical Report
%
% Author: [Your Name]
% Institution: [Your Institution]
% Date: December 2025
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt,a4paper]{report}

%------------------------------------------------------------------------------
% Packages
%------------------------------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{mathtools}
\usepackage{physics}
\usepackage{braket}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{quantikz}
\usepackage[sorting=none,style=numeric-comp]{biblatex}
\usepackage{geometry}
\usepackage{tcolorbox}

\geometry{margin=2.5cm}
\graphicspath{{figures/}}

%------------------------------------------------------------------------------
% Custom Commands
%------------------------------------------------------------------------------
\newcommand{\Nc}{N_c}
\newcommand{\Nq}{n_q}
\newcommand{\Ng}{n_g}
\newcommand{\SU}[1]{\mathrm{SU}(#1)}
\renewcommand{\Tr}{\operatorname{Tr}}
\newcommand{\CF}{C_F}
\newcommand{\CA}{C_A}
\newcommand{\lhat}[1]{\hat{\lambda}_{#1}}
\newcommand{\gellmann}[1]{\lambda_{#1}}
\newcommand{\generator}[1]{T^{#1}}
\newcommand{\structconst}[3]{f^{#1 #2 #3}}

%------------------------------------------------------------------------------
% Theorem Environments
%------------------------------------------------------------------------------
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{remark}{Remark}[chapter]
\newtheorem{example}{Example}[chapter]

%------------------------------------------------------------------------------
% Code Listings Style
%------------------------------------------------------------------------------
\lstset{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\bfseries,
    commentstyle=\color{gray}\itshape,
    stringstyle=\color{red},
    showstringspaces=false,
    breaklines=true,
    frame=single,
    numbers=left,
    numberstyle=\tiny\color{gray},
    captionpos=b,
}

%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
% \addbibresource{references.bib}

%------------------------------------------------------------------------------
% Document Information
%------------------------------------------------------------------------------
\title{%
    \textbf{Quantum Computation of QCD Colour Factors} \\[1em]
    \Large Implementation and Verification of the Chawdhry--Pellen Algorithm \\[0.5em]
    \large Using the QC-Amp Software Library
}
\author{%
    [Your Name] \\
    \small [Department] \\
    \small [University] \\
    \small \texttt{[email@university.edu]}
}
\date{December 2025}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

\begin{abstract}
\noindent
We present \texttt{QC-Amp}, a Python software library implementing the quantum algorithm of Chawdhry and Pellen for computing colour factors in Quantum Chromodynamics (QCD) scattering amplitudes. The library provides a complete implementation of the unitarisation scheme required to encode the non-unitary SU(3) generators into quantum circuits, along with verification against analytically known results. We demonstrate the successful computation of the colour factor $C = 4$ for the quark self-energy diagram, achieving exact agreement with the theoretical prediction. This work establishes a foundation for extending quantum amplitude calculations to more complex multi-parton processes relevant to high-energy physics phenomenology at the Large Hadron Collider and future colliders.

\vspace{1em}
\noindent\textbf{Keywords:} Quantum Computing, Quantum Chromodynamics, Colour Factors, Scattering Amplitudes, Qiskit, SU(3)
\end{abstract}

\tableofcontents

%------------------------------------------------------------------------------
\chapter{Introduction}
\label{ch:introduction}
%------------------------------------------------------------------------------

\input{chapters/introduction}

%------------------------------------------------------------------------------
\chapter{Theoretical Background}
\label{ch:theory}
%------------------------------------------------------------------------------

\input{chapters/theory}

%------------------------------------------------------------------------------
\chapter{The Chawdhry--Pellen Algorithm}
\label{ch:algorithm}
%------------------------------------------------------------------------------

\input{chapters/algorithm}

%------------------------------------------------------------------------------
\chapter{Implementation}
\label{ch:implementation}
%------------------------------------------------------------------------------

\input{chapters/implementation}

%------------------------------------------------------------------------------
\chapter{Results and Verification}
\label{ch:results}
%------------------------------------------------------------------------------

\input{chapters/results}

%------------------------------------------------------------------------------
\chapter{Conclusions and Future Work}
\label{ch:conclusions}
%------------------------------------------------------------------------------

\input{chapters/conclusions}

%------------------------------------------------------------------------------
% Appendices
%------------------------------------------------------------------------------
\appendix

\chapter{Mathematical Details, Code Listings, and Additional Results}
\label{app:appendices}

\input{chapters/appendices}

%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
% \printbibliography[heading=bibintoc,title={References}]

% Temporary manual bibliography until biblatex is configured
% Full bibliography in references.bib
\begin{thebibliography}{99}

\bibitem{ChawdhryPellen2023}
H.~A.~Chawdhry and M.~Pellen,
``Quantum simulation of colour in perturbative quantum chromodynamics,''
\textit{Phys. Rev. D}\ \textbf{108}, 094028 (2023),
\href{https://arxiv.org/abs/2303.04818}{arXiv:2303.04818 [hep-ph]}.

\bibitem{Feynman1982}
R.~P.~Feynman,
``Simulating physics with computers,''
\textit{Int. J. Theor. Phys.}\ \textbf{21}, 467--488 (1982).

\bibitem{Nielsen2010}
M.~A.~Nielsen and I.~L.~Chuang,
\textit{Quantum Computation and Quantum Information},
10th Anniversary Edition (Cambridge University Press, 2010).

\bibitem{Peskin1995}
M.~E.~Peskin and D.~V.~Schroeder,
\textit{An Introduction to Quantum Field Theory}
(Westview Press, 1995).

\bibitem{Ellis1996}
R.~K.~Ellis, W.~J.~Stirling, and B.~R.~Webber,
\textit{QCD and Collider Physics}
(Cambridge University Press, 1996).

\bibitem{Muta2010}
T.~Muta,
\textit{Foundations of Quantum Chromodynamics},
3rd Edition (World Scientific, 2010).

\bibitem{Georgi1999}
H.~Georgi,
\textit{Lie Algebras in Particle Physics},
2nd Edition (Perseus Books, 1999).

\bibitem{Qiskit2024}
Qiskit contributors,
``Qiskit: An Open-source Framework for Quantum Computing,''
(2024), \url{https://qiskit.org/}.

\bibitem{GellMann1962}
M.~Gell-Mann,
``Symmetries of Baryons and Mesons,''
\textit{Phys. Rev.}\ \textbf{125}, 1067--1084 (1962).

\bibitem{Gross1973}
D.~J.~Gross and F.~Wilczek,
``Ultraviolet Behavior of Non-Abelian Gauge Theories,''
\textit{Phys. Rev. Lett.}\ \textbf{30}, 1343--1346 (1973).

\bibitem{Politzer1973}
H.~D.~Politzer,
``Reliable Perturbative Results for Strong Interactions?''
\textit{Phys. Rev. Lett.}\ \textbf{30}, 1346--1349 (1973).

\bibitem{Jordan2012}
S.~P.~Jordan, K.~S.~M.~Lee, and J.~Preskill,
``Quantum Algorithms for Quantum Field Theories,''
\textit{Science}\ \textbf{336}, 1130--1133 (2012).

\bibitem{Preskill2018}
J.~Preskill,
``Quantum Computing in the NISQ era and beyond,''
\textit{Quantum}\ \textbf{2}, 79 (2018).

\bibitem{Bauer2023}
C.~W.~Bauer \textit{et al.},
``Quantum Simulation for High Energy Physics,''
\textit{PRX Quantum}\ \textbf{4}, 027001 (2023).

\bibitem{Sjodahl2012}
M.~Sj{\"o}dahl,
``ColorMath -- A package for color summed calculations in SU(Nc),''
\textit{Eur. Phys. J. C}\ \textbf{73}, 2310 (2013).

\bibitem{vanRitbergen1999}
T.~van~Ritbergen, A.~N.~Schellekens, and J.~A.~M.~Vermaseren,
``Group theory factors for Feynman diagrams,''
\textit{Int. J. Mod. Phys. A}\ \textbf{14}, 41--96 (1999).

\bibitem{numpy2020}
C.~R.~Harris \textit{et al.},
``Array programming with NumPy,''
\textit{Nature}\ \textbf{585}, 357--362 (2020).

\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
